---
title: "Hnf4a Manuscript Figure 5"
output: github_document
---

Loading packages

```{r}
#Loading required libraries
library(ggplot2)
library(plyr)
library(tidyverse)
library(phyloseq) #To open phyloseq objects output by Hardac
library(RColorBrewer) 
library(pals) #Improved color palette options
library(ggpubr) #Improved ggplot functions
library(vegan)
library(DESeq2)
```

Alpha Diversity


```{r}

ps2 <- read_rds("phyloseq_v2.rds")
ps2
```

What does the sample_data look like?

```{r}
df1 <- ps2@sam_data
head(df1)
```

First, we need to group HET and WT mice to compare alpha and beta div between WT and KO mice


```{r}
#Removing negative controls
ps2.noctrl = subset_samples(ps2, !is.na(mouse_number))
ps2.noctrl
```

```{r}
#Removing positive controls
ps2.noctrlv2 = subset_samples(ps2.noctrl, note != "positive control")
ps2.noctrlv2

```

```{r}
#Copying the genotype variable
ps2.noctrlv2@sam_data$Genotypev2 <- ps2.noctrlv2@sam_data$Genotype
ps2.noctrlv2

```

```{r}
ps2.noctrlv2@sam_data$Genotypev2 <- gsub("HET", "WT", ps2.noctrlv2@sam_data$Genotypev2)
```

Let us look at the new sample data header

```{r}
df2 <- ps2.noctrlv2@sam_data
head(df2)
```

```{r}
#Let us confirm that all HETs were converted to WT
 df2[["Genotypev2"]]
```

```{r}
df1[["Genotype"]]
```

Data preprocessing for ordination. Rare and low abundance taxa can swamp out major trends (E.g. Genotype)

```{r}
sample_min_count = 50

ps2 %>%
  prune_samples(sample_sums(.)>=sample_min_count, .) ->
  ps2.sample_prune

```

```{r}
min_count = 3
min_sample_frac = 0.10

prune.vec = filter_taxa(ps2.sample_prune, 
                       function(x) sum(x >= min_count) >= (min_sample_frac*length(x)))
ps2.st_prune = prune_taxa(prune.vec, ps2.sample_prune)
ntaxa(ps2.st_prune)

```

```{r}
ps2.st_prune.even = transform_sample_counts(ps2.st_prune, function(x) 1E6 * x/sum(x))

```

```{r}
alpha_measures_all <- c("Shannon", "Chao1", "Simpson")
ga1 <- plot_richness(ps2.noctrlv2, measures = alpha_measures_all, x ="Genotypev2", color = "Genotypev2") + geom_boxplot() +theme_minimal() +scale_color_brewer(palette = "Set1")
ga1
```

Are the differences in alpha div significant?

```{r}
erich <- estimate_richness(ps2.noctrlv2, measures = c("Chao1","Shannon","Simpson"))
ttest1 <- t(sapply(erich, function(x) unlist(kruskal.test(x~sample_data(ps2.noctrlv2)$Genotypev2)[c("estimate","p.value","statistic","conf.int")])))
ttest1
```

Repeating for inflamed vs. non-inflamed mice

```{r}
erich <- estimate_richness(ps2.noctrlv2, measures = c("Chao1","Shannon","Simpson"))
ttest1 <- t(sapply(erich, function(x) unlist(kruskal.test(x~sample_data(ps2.noctrlv2)$inflamed)[c("estimate","p.value","statistic","conf.int")])))
ttest1
```

Beta Diversity


```{r}
bc_dist = phyloseq::distance(ps2.noctrlv2, method="bray", weighted=F)
adonis2(bc_dist ~ sample_data(ps2.noctrlv2)$Genotypev2)
```

Repeating for inflamed vs. non-inflamed mice

```{r}
bc_dist = phyloseq::distance(ps2.noctrlv2, method="bray", weighted=F)
adonis2(bc_dist ~ sample_data(ps2.noctrlv2)$inflamed)
```


Panel 5A: PCoA plot colored according to genotype

```{r}
ps2.st_prune.even
```

Log transforming lipocalin (Lcn2) values


```{r}
ps2.st_prune.even@sam_data$lipocalin_log <- log10(ps2.st_prune.even@sam_data$lipocalin_ng)
```

```{r}
ps2.st_prune.even = subset_samples(ps2.st_prune.even, !is.na(mouse_number))
ps2.st_prune.even
```

```{r}
ps2.st_prune.even = subset_samples(ps2.st_prune.even, note != "positive control")
ps2.st_prune.even
```


```{r}
ps2.st_prune.even.pcoa_bc <- ordinate(ps2.st_prune.even, "PCoA", "bray")
p3test <- plot_ordination(ps2.st_prune.even, ps2.st_prune.even.pcoa_bc, type="samples", color="Genotype") + 
  stat_ellipse(type = "norm") +
  theme_bw() + scale_color_manual(values = c("#040404", "#ec3525", "#445ca4"))
p3test

```

Panel 5B: PCoA plot colored according to log

```{r}
p7testlog <- plot_ordination(ps2.st_prune.even, ps2.st_prune.even.pcoa_bc, type="samples", color = "lipocalin_log", shape = "Genotype") +
  scale_color_gradient(name = NULL,
                       high = "#eae82e",
                       low = "#2c2f81") +
stat_ellipse(type = "norm") +
  theme_bw()
p7testlog
```


Panel 5C: Relative Abundance plot in the same order as the lipocalin heatmap


```{r}
#Should add up to 64 samples total, with controls removed it should be 52

ps2.pct = transform_sample_counts(ps2, function(x) 100 * x/sum(x))
ps2.pct

ps2.pct.noctrl1 = subset_samples(ps2.pct, !is.na(new_mouse_number_for_panel_C))
ps2.pct.noctrl1

ps2.pct.noctrl = subset_samples(ps2.pct.noctrl1, note == "")
ps2.pct.noctrl


```

```{r}
top25 <- names(sort(taxa_sums(ps2.pct.noctrl), decreasing=TRUE))[1:25]
ps.top25 <- prune_taxa(top25, ps2.pct.noctrl)
barplot2.2 <- plot_bar(ps.top25, x=as.character('new_mouse_number_for_panel_C'), fill = "Genus") + 
  geom_bar(stat = "identity", position = "stack", size=0)+ scale_fill_manual(values = as.vector(watlington())) + coord_flip() + theme_bw() +theme(legend.position = "bottom", legend.box = "horizontal")
barplot2.2
```


```{r}
barplot2.2$data$new_mouse_number_for_panel_C <- as.factor(barplot2.2$data$new_mouse_number_for_panel_C)
```

```{r}
barplot2.2 + scale_x_discrete(limits=rev)
```


Deseq2

```{r}
hnf4a_dds <- phyloseq_to_deseq2(ps2.noctrlv2, ~ Genotypev2)

```


```{r}
hnf4a_dds
```


```{r}
# First we remove rows with very few counts E.g. 5
hnf4a_dds <- hnf4a_dds[ rowSums(counts(hnf4a_dds)) > 5, ]
```


```{r}
cts_hnf4a <- counts(hnf4a_dds)
geoMeans_hnf4a <- apply(cts_hnf4a, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds_hnf4a <- estimateSizeFactors(hnf4a_dds, geoMeans=geoMeans_hnf4a)
```



```{r}
dds_hnf4a <- DESeq(dds_hnf4a, test = "Wald", fitType = "parametric")

```

```{r}
#Using the results table in Deseq2 to create a table with significant results for plotting and viewing this table to ensure it is properly formatted
res_hnf4a = results(dds_hnf4a, cooksCutoff = FALSE)
alpha = 0.01
sigtab_hnf4a = res_hnf4a[which(res_hnf4a$padj < alpha), ]
sigtab_hnf4a = cbind(as(sigtab_hnf4a, "data.frame"), as(tax_table(ps2.noctrlv2)[rownames(sigtab_hnf4a), ], "matrix"))
head(sigtab_hnf4a)

```


```{r}
#Saving Deseq object as tab separated values
#write.table(sigtab_hnf4a, file = "Deseq2_hnf4a-flaring_WT-HET-vs-KO_ASV-names.tsv", row.names=TRUE, sep="\t")

```


```{r}
#Which comparison is being made? This code shows that it is genotype "wt" vs "ko"
#The second line of code dim(sigtab) shows number of differential taxa at the chosen alpha
resultsNames(dds_hnf4a)
dim(sigtab_hnf4a)
```


Repeating the above for inflamed vs. non-inflamed KO mice

First, we need to remove all WT (Genotypev2) samples so we can compare just the inflamed and non-inflamed KO mice

```{r}
#Removing WT (Includes HET for Genotype2 variable)
ps2.noctrlv2_noWT = subset_samples(ps2.noctrlv2, Genotypev2 != "WT")
ps2.noctrlv2_noWT

```

```{r}
hnf4a_noWT_dds <- phyloseq_to_deseq2(ps2.noctrlv2_noWT, ~ inflamed)

```


```{r}
hnf4a_noWT_dds
```


```{r}
# First we remove rows with very few counts E.g. 5
hnf4a_noWT_dds <- hnf4a_noWT_dds[ rowSums(counts(hnf4a_noWT_dds)) > 5, ]
```


```{r}
cts_hnf4a_noWT <- counts(hnf4a_noWT_dds)
geoMeans_hnf4a_noWT <- apply(cts_hnf4a_noWT, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds_hnf4a_noWT <- estimateSizeFactors(hnf4a_noWT_dds, geoMeans=geoMeans_hnf4a_noWT)
```



```{r}
dds_hnf4a_noWT <- DESeq(dds_hnf4a_noWT, test = "Wald", fitType = "parametric")

```

```{r}
#Using the results table in Deseq2 to create a table with significant results for plotting and viewing this table to ensure it is properly formatted
res_hnf4a_noWT = results(dds_hnf4a_noWT, cooksCutoff = FALSE)
alpha = 0.01
sigtab_hnf4a_noWT = res_hnf4a_noWT[which(res_hnf4a_noWT$padj < alpha), ]
sigtab_hnf4a_noWT = cbind(as(sigtab_hnf4a_noWT, "data.frame"), as(tax_table(ps2.noctrlv2_noWT)[rownames(sigtab_hnf4a_noWT), ], "matrix"))
head(sigtab_hnf4a_noWT)

```


```{r}
#Saving Deseq object as tab separated values
#write.table(sigtab_hnf4a_noWT, file = "Deseq2_hnf4a_noWT-inflamed-ko_Y-v-N_ASV-names.tsv", row.names=TRUE, sep="\t")

```


```{r}
#Which comparison is being made? This code shows that it is genotype "wt" vs "ko"
#The second line of code dim(sigtab) shows number of differential taxa at the chosen alpha
resultsNames(dds_hnf4a_noWT)
dim(sigtab_hnf4a_noWT)
```


```{r}
sessionInfo()
```



