---
title: "Hnf4a Manuscript Figure 5"
output: html_notebook
---

Loading packages

```{r}
#Loading required libraries
library(ggplot2)
library(plyr)
library(tidyverse)
library(phyloseq) #To open phyloseq objects output by Hardac
library(RColorBrewer) 
library(pals) #Improved color palette options
library(ggpubr) #Improved ggplot functions
library(vegan)
```

Alpha Diversity


```{r}
#Opening phyloseq object without the Hearts and Parks (HP) samples. There should be 804 samples total

ps2 <- read_rds("/Users/jay/Library/CloudStorage/Box-Box/Hardac output/Cece hnf4a flaring 5-21/phyloseq.rds")
ps2
```

What does the sample_data look like?

```{r}
df1 <- ps2@sam_data
head(df1)
```

First, we need to group HET and WT mice to compare alpha and beta div between WT and KO mice


```{r}
#Removing negative controls
ps2.noctrl = subset_samples(ps2, !is.na(mouse_number))
ps2.noctrl
```

```{r}
#Removing positive controls
ps2.noctrlv2 = subset_samples(ps2.noctrl, note != "positive control")
ps2.noctrlv2

```

```{r}
#Copying the genotype variable
ps2.noctrlv2@sam_data$Genotypev2 <- ps2.noctrlv2@sam_data$Genotype
ps2.noctrlv2

```

```{r}
ps2.noctrlv2@sam_data$Genotypev2 <- gsub("HET", "WT", ps2.noctrlv2@sam_data$Genotypev2)
```

Let us look at the new sample data header

```{r}
df2 <- ps2.noctrlv2@sam_data
head(df2)
```

```{r}
#Let us confirm that all HETs were converted to WT
 df2[["Genotypev2"]]
```

```{r}
df1[["Genotype"]]
```

Data preprocessing for ordination. Rare and low abundance taxa can swamp out major trends (E.g. Genotype)

```{r}
sample_min_count = 50

ps2 %>%
  prune_samples(sample_sums(.)>=sample_min_count, .) ->
  ps2.sample_prune

```

```{r}
min_count = 3
min_sample_frac = 0.10

prune.vec = filter_taxa(ps2.sample_prune, 
                       function(x) sum(x >= min_count) >= (min_sample_frac*length(x)))
ps2.st_prune = prune_taxa(prune.vec, ps2.sample_prune)
ntaxa(ps2.st_prune)

```

```{r}
ps2.st_prune.even = transform_sample_counts(ps2.st_prune, function(x) 1E6 * x/sum(x))

```

```{r}
alpha_measures_all <- c("Shannon", "Chao1", "Simpson")
ga1 <- plot_richness(ps2.noctrlv2, measures = alpha_measures_all, x ="Genotypev2", color = "Genotypev2") + geom_boxplot() +theme_minimal() +scale_color_brewer(palette = "Set1")
ga1
```

Are the differences in alpha div significant?

```{r}
erich <- estimate_richness(ps2.noctrlv2, measures = c("Chao1","Shannon","Simpson"))
ttest1 <- t(sapply(erich, function(x) unlist(kruskal.test(x~sample_data(ps2.noctrlv2)$Genotypev2)[c("estimate","p.value","statistic","conf.int")])))
ttest1
```

Repeating for inflamed vs. non-inflamed mice

```{r}
erich <- estimate_richness(ps2.noctrlv2, measures = c("Chao1","Shannon","Simpson"))
ttest1 <- t(sapply(erich, function(x) unlist(kruskal.test(x~sample_data(ps2.noctrlv2)$inflamed)[c("estimate","p.value","statistic","conf.int")])))
ttest1
```

Beta Diversity


```{r}
bc_dist = phyloseq::distance(ps2.noctrlv2, method="bray", weighted=F)
adonis2(bc_dist ~ sample_data(ps2.noctrlv2)$Genotypev2)
```

Repeating for inflamed vs. non-inflamed mice

```{r}
bc_dist = phyloseq::distance(ps2.noctrlv2, method="bray", weighted=F)
adonis2(bc_dist ~ sample_data(ps2.noctrlv2)$inflamed)
```


Panel 5A: PCoA plot colored according to genotype

```{r}
ps2.st_prune.even
```

Log transforming lipocalin (Lcn2) values


```{r}
ps2.st_prune.even@sam_data$lipocalin_log <- log10(ps2.st_prune.even@sam_data$lipocalin_ng)
```

```{r}
ps2.st_prune.even = subset_samples(ps2.st_prune.even, !is.na(mouse_number))
ps2.st_prune.even
```

```{r}
ps2.st_prune.even = subset_samples(ps2.st_prune.even, note != "positive control")
ps2.st_prune.even
```


```{r}
ps2.st_prune.even.pcoa_bc <- ordinate(ps2.st_prune.even, "PCoA", "bray")
p3test <- plot_ordination(ps2.st_prune.even, ps2.st_prune.even.pcoa_bc, type="samples", color="Genotype") + 
  stat_ellipse(type = "norm") +
  theme_bw() + scale_color_manual(values = c("#040404", "#ec3525", "#445ca4"))
p3test

```

Panel 5B: PCoA plot colored according to log

```{r}
p7testlog <- plot_ordination(ps2.st_prune.even, ps2.st_prune.even.pcoa_bc, type="samples", color = "lipocalin_log", shape = "Genotype") +
  scale_color_gradient(name = NULL,
                       high = "#eae82e",
                       low = "#2c2f81") +
stat_ellipse(type = "norm") +
  theme_bw()
p7testlog
```





Panel 5C



Deseq2




```{r}
sessionInfo()
```



