---
title: "Hnf4a Manuscript Figure 6"
output: github_document
---


Loading packages

```{r}
#Loading required libraries
library(ggplot2)
library(plyr)
library(tidyverse)
library(phyloseq) #To open phyloseq objects output by Hardac
library(RColorBrewer) 
library(pals) #Improved color palette options
library(ggpubr) #Improved ggplot functions
library(vegan)
library(DESeq2)
```

Panel 6A PCoA plot


```{r}
ps2 <- read_rds("phyloseq_lipocalin_HET-WT.rds")
ps2
```

Panel 6B PCoA plot Axis a over time with GLM statistics

The N/A sample doesn't add much to the analysis, removing this for the rest of the session

```{r}
ps2 = subset_samples(ps2, genotype != "N/A")
```



Data preprocessing for ordination. Rare and low abundance taxa can swamp out major trends (E.g. Genotype)

```{r}
sample_min_count = 50

ps2 %>%
  prune_samples(sample_sums(.)>=sample_min_count, .) ->
  ps2.sample_prune

```

```{r}
min_count = 3
min_sample_frac = 0.10

prune.vec = filter_taxa(ps2.sample_prune, 
                       function(x) sum(x >= min_count) >= (min_sample_frac*length(x)))
ps2.st_prune = prune_taxa(prune.vec, ps2.sample_prune)
ntaxa(ps2.st_prune)

```

```{r}
ps2.st_prune.even = transform_sample_counts(ps2.st_prune, function(x) 1E6 * x/sum(x))

```

```{r}
ps2.st_prune.even.pcoa_bc <- ordinate(ps2.st_prune.even, "PCoA", "bray")
pltest <- plot_ordination(ps2.st_prune.even, ps2.st_prune.even.pcoa_bc, type="samples", color="genotype")
pltest

```

```{r}
#This changes the variable "age" to numeric
class(sample_data(ps2.st_prune.even)[["age"]]) <- "numeric"
```

```{r}
p7 <- plot_ordination(ps2.st_prune.even, ps2.st_prune.even.pcoa_bc, type="samples", color="age", shape = "genotype") +
  scale_color_gradient(name = NULL,
                       high = "#FF8000",
                       low = "#FFE5CC")+ theme_bw()
p7
```

Plotting just PC Axis 1 stacked over time

```{r}
df_pcoa <- p7$data
head(df_pcoa)
```

Example taken from this link: http://www.sthda.com/english/wiki/ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization



```{r}
p2 <- ggplot(df_pcoa, aes(x=age, y=Axis.1, color=genotype)) + geom_point() +theme_bw()

p2
```

```{r}

df_pcoa$age <- as.factor(df_pcoa$age)
sapply(df_pcoa, class)
```

```{r}
model_pcoa <- glm(df_pcoa$Axis.1 ~ genotype2 + age, data = df_pcoa)
```


```{r}
summary(model_pcoa)$coef
```

```{r}
#Writing to dataframe 
df_pcoa_genotype2 <- as.data.frame(summary(model_pcoa)$coef)
head(df_pcoa_genotype2)
```


DESeq2 analysis of WT+HET vs KO

```{r}
hnfdds2 <- phyloseq_to_deseq2(ps2, ~ genotype2)
hnfdds2
```


```{r}
# First we remove rows with very few counts E.g. 5
hnfdds2 <- hnfdds2[ rowSums(counts(hnfdds2)) > 5, ]
```

```{r}
cts <- counts(hnfdds2)
geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
dds3 <- estimateSizeFactors(hnfdds2, geoMeans=geoMeans)
```

```{r}
dds3 <- DESeq(dds3, test = "Wald", fitType = "parametric")

```

```{r}
resultsNames(dds3)
```



```{r}
#Using the results table in Deseq2 to create a table with significant results for plotting and viewing this table to ensure it is properly formatted
res = results(dds3, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps2)[rownames(sigtab), ], "matrix"))
head(sigtab)
```

```{r}
dim(sigtab)
```


```{r}
#Saving lipocalin flaring Deseq object as tab separated values
#write.table(sigtab, file = "Deseq2_HET-WT_vs_KO_6-18-2020_V2.tsv", row.names=TRUE, sep="\t")

```


```{r}
#Which comparison is being made? This code shows that it is flare condition "Yes" vs "No"
#The second line of code dim(sigtab) shows that there are 56 differential taxa
resultsNames(dds3)
dim(sigtab)
```

```{r}
#Just coloring by species, 6-18-2020, repeating this with the genotype 2 label
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ds2 <- ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Species)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
ds2 + ggtitle("Log2fold change in WT+HET vs. KO")

```


ALDEx2 correlation of Log2(Lcn2) vs. CLR transformed ASV counts


Correlation (Spearman) of Delta Log2(Lcn2) vs. Delta ASVs


Panel 6C Relative Abundance of top 20 ASV for each mouse over 52 Weeks



Alpha diversity




